Partridge: An effective system for the automatic classification of the types of academic papers

### 2013-05-23 doc/bcspaper/article.tex
Maria intro edits

% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass{svmult}
%
%% choose options for [] as required from the list
%% in the Reference Guide
%
\usepackage{mathptmx}       % selects Times Roman as basic font
\usepackage{helvet}         % selects Helvetica as sans-serif font
\usepackage{courier}        % selects Courier as typewriter font
\usepackage{type1cm}        % activate if the above 3 fonts are
                             % not available on your system
%
\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
%                             % when including figure files
\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom
\usepackage{subcaption}
\captionsetup{compatibility=false}
%
%% see the list of further useful packages
%% in the Reference Guide
%
%\makeindex             % used for the subject index
%                       % please use the style svind.ist with
%                       % your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\begin{document}

@@

\abstract{Partridge is a system that enables intelligent search for academic
papers. Here we describe how Partridge implements the automatic classification
of the type of a publication. For each paper, Partridge automatically extracts
the full paper content from PDF files, determines sentence boundaries,
automatically labels the sentences with core scientific concepts, and then uses
a random forest model to classify the paper type. We show that the type of a
paper can be reliably predicted by a model which analyses the distribution of
core scientific concepts within the sentences of the paper. We discuss the
appropriateness of many of the existing paper types used by major journals, and
their corresponding distributions. Partridge is online and available for use,
includes a browser-friendly bookmarklet for new paper submission, and
demonstrates a range of possibilities for more intelligent search in the
scientific literature. The Partidge instance can be used at
\url{http://farnsworth.papro.org.uk}, and further information about the project
can be found at \url{http://papro.org.uk}}

\section{Introduction} \label{sec:1} Since the advent of the `Digital Age' and
the ability of computers to copy and reproduce information for a negligible
cost, the amount of information available to researchers has been increasing
drastically.  B-C Bj\"{o}rk (2009) estimates that approximately 1.4 Million
papers were published in the year 2006 alone\cite{bjork2009}.  As available
information increases, relevant material becomes progressively more difficult
to find manually and the need for an automated information retrieval tool more
apparent.  There are already a large number of information retrieval and
recommendation systems for scientific publications.  Many of these systems,
such as AGRICOLA (\url{http://agricola.nal.usda.gov/}), the Cochrane
Library(\url{http://www.thecochranelibrary.com/}) and Textpresso
(\url{http://www.textpresso.org/}) index only publications from predefined
journals or topics (for the above examples, Agriculture, Biology and
Bioinformatics respectively).  Unfortunately, these domain specific indexing
systems usually only contain a small subset of papers, excluding potentially
crucial literature because it does not quite fit into the subject domain. 
%This problem is often exacerbated by the decision to leave out some papers and
%articles because the system administrators do not have permission from the
%author or publisher to include them. 
The value of these systems to their users is often restricted by the small
proportion of available literature that they index, forcing researchers to use
multiple, domain specific, search engines for their queries.  In contrast,
there are also a number of interdisciplinary indexing systems and online
journals such as arXiv( \url{http://arxiv.org/}),
PloSOne(\url{http://plosone.org/}), and JSTOR (\url{http://www.jstor.org/}),
that try to incorporate wide ranges of papers from as many disciplines as
possible. The traits of these systems often complement those of their
domain-specific counterparts; they provide a comprehensive collection of
literature but insufficient filtering and indexing capabilities.  One of the
most publicised and well known interdisciplinery scientific literature search
systems is Google Scholar (\url{http://scholar.google.com}). Google offers
advanced query options specific to Scholar that allow searching by author, year
and for words that occur only in the document title.

However, the document title is just one part of the important structure to be
found within a scientific document.  Liakata \emph{et al.} (2012) describe a
system for automatically processing and classifying sentences in a research
paper according to the core scientific concept (CoreSC) that they
describe\cite{Liakata2012}.  There are 13 CoreSCs, including {\em hypothesis},
{\it background}, {\em result}, {\em method} and {\em conclusion}.  CoreSC
labels can be allocated to all sentences in a scientific paper in order to
identify which scientific concept each sentence encapsulates.  SAPIENTA
(\url{http://www.sapientaproject.com}) is a machine learning application which
can automatically annotate all sentences in a scientific paper with their
CoreSC labels. It was trained using a corpus of physical chemistry and
biochemistry research papers whose sentences were manually annotated using the
CoreSC\cite{LIAKATA10.644} scheme.  An intelligent academic information
retrieval system can use this information in order to provide better filtering
and search capabilities for researchers.  The ability to search for phrases and
keywords by CoreSC will facilitate context-aware keyword search, that allows
researchers to only accept papers where a term appears in sentences with a
specific CoreSC label.  This can be used to greatly improve both the precision
with which researchers are able to perform searches for scientific literature
and the accuracy of those searches in terms of relevance to the reader.

@@

This model of paper type is currently in use in our Partridge system, which has
been created as an intelligent full-text search platform for scientific papers.
Partridge can make use of automatically derived CoreSC sentence labels and
automatically derived paper types, to allow deeper information queries.  We
discuss the reliability of this model of paper types and the insights that have
been gained for the authorship of papers.

>>>>

% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass{svmult}
%
%% choose options for [] as required from the list
%% in the Reference Guide
%
\usepackage{mathptmx}       % selects Times Roman as basic font
\usepackage{helvet}         % selects Helvetica as sans-serif font
\usepackage{courier}        % selects Courier as typewriter font
\usepackage{type1cm}        % activate if the above 3 fonts are
                             % not available on your system
%
\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
%                             % when including figure files
\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom
\usepackage{subcaption}
\usepackage{soul}
\captionsetup{compatibility=false}
%
%% see the list of further useful packages
%% in the Reference Guide
%
%\makeindex             % used for the subject index
%                       % please use the style svind.ist with
%                       % your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\begin{document}

@@

\abstract{Partridge is a system that enables intelligent search for
  academic papers by allowing users to query particular terms within
  sentences designating a particular core scientific concept (e.g. Hypotheses, Results, etc). 
  The system also automatically classifies papers according to
  article types (e.g. Review, Case Study). Here we focus on the latter
  aspect of the system. For each paper, Partridge automatically
  extracts the full paper content from PDF files, converts it to XML,
  determines sentence boundaries, automatically labels the sentences
  with core scientific concepts, and then uses a random forest model to
  classify the paper type. We show that the type of a paper can be
  reliably predicted by a model which analyses the distribution of
  core scientific concepts within the sentences of the paper. We
  discuss the appropriateness of many of the existing paper types used
  by major journals, and their corresponding distributions. Partridge
  is online and available for use, includes a browser-friendly
  bookmarklet for new paper submission, and demonstrates a range of
  possibilities for more intelligent search in the scientific
  literature. The Partidge instance can be used at
  \url{http://farnsworth.papro.org.uk}, and further information about
  the project can be found at \url{http://papro.org.uk}}

\section{Introduction} \label{sec:1} Since the advent of the `Digital Age' and
the ability of computers to copy and reproduce information for a negligible
cost, the amount of information available to researchers has been increasing
drastically. \st{ B-C Bj\"{o}rk (2009) estimates that approximately 1.4 Million
papers were published in the year 2006 alone\cite{bjork2009}}. As available
information increases, relevant material becomes progressively more difficult
to find manually and the need for an automated information retrieval tool more
apparent.  There are already a large number of information retrieval and
recommendation systems for scientific publications.  Many of these systems,
such as AGRICOLA (\url{http://agricola.nal.usda.gov/}), the Cochrane
Library(\url{http://www.thecochranelibrary.com/}) and Textpresso
(\url{http://www.textpresso.org/}) index only publications from predefined
journals or topics (for the above examples, Agriculture, Biology and
Bioinformatics respectively).  Unfortunately, these domain specific indexing
systems usually only contain a small subset of papers, excluding potentially
crucial literature because it does not quite fit into the subject domain. 
%This problem is often exacerbated by the decision to leave out some papers and
%articles because the system administrators do not have permission from the
%author or publisher to include them. 
The value of these systems to their users is often restricted by the small
proportion of available literature that they index, forcing researchers to use
multiple, domain specific, search engines for their queries.  In contrast,
there are also a number of interdisciplinary indexing systems and online
journals such as arXiv( \url{http://arxiv.org/}),
PloSOne(\url{http://plosone.org/}), and JSTOR (\url{http://www.jstor.org/}),
that try to incorporate wide ranges of papers from as many disciplines as
possible. The traits of these systems often complement those of their
domain-specific counterparts; they provide a comprehensive collection of
literature but insufficient filtering and indexing capabilities.  One of the
most publicised and well known interdisciplinery scientific literature search
systems is Google Scholar (\url{http://scholar.google.com}). Google offers
advanced query options specific to Scholar that allow searching by author, year
and for words that occur only in the document title.

However, the document title is just one part of the important structure to be
found within a scientific document.  Liakata \emph{et al.} (2012) describe a
system for automatically processing and classifying sentences in a research
paper according to the core scientific concept (CoreSC) that they
describe\cite{Liakata2012}.  There are 11 CoreSCs, including {\em Hypothesis}, {\em Goal},
{\it Background}, {\em Method}, {\em Result} and {\em Conclusion}.  CoreSC
labels can be allocated to all sentences in a scientific paper in order to
identify which scientific concept each sentence encapsulates.  SAPIENTA
(\url{http://www.sapientaproject.com}) is a publicly available machine learning application which
can automatically annotate all sentences in a scientific paper with their
CoreSC labels. It was trained using a corpus of physical chemistry and
biochemistry research papers whose sentences were manually annotated using the
CoreSC\cite{LIAKATA10.644} scheme.  An intelligent academic information
retrieval system can use this information in order to provide better filtering
and search capabilities for researchers.  Partridge {\st the ability to search for phrases and
keywords by CoreSC will facilitate} implements such context-aware keyword search, by allowing
researchers to search for papers where a term appears in sentences with a
specific CoreSC label (e.g only in {\em Method} senetences). This can be used to greatly improve both the precision
with which researchers are able to perform searches for scientific literature
and the accuracy of those searches in terms of relevance to the reader.

@@

This model of paper type is currently in use in our Partridge system, which has
been created as an intelligent full-text search platform for scientific papers.
Partridge makes use of automatically derived CoreSC sentence labels and
automatically derived paper types, to allow deeper information queries.  We
discuss the reliability of this model of paper types and the insights that have
been gained for the authorship of papers.
